### 2016-03-24 ch02-qlogms/excitation-design.tex
More details on computation Fi

In open loop experiments, the scaled average information matrix is given by
\begin{align}
\IF & = \lim_{N\rightarrow\infty}
          \frac{1}{N}
            \sum_{t=1}^{N}
              {\left(\frac{\partial\hat{y}(t)}{\partial\theta}\right)
               \left(\frac{\partial\hat{y}(t)}{\partial\theta}\right)^{\T}},\\
\frac{\partial\hat{y}(t)}{\partial\theta}
    & =\frac{\partial G(s,\theta)}{\partial \theta}u(t).
\end{align}
By Parseval's theorem, $\IF$ can be expressed as
\begin{align}
\IF &= \frac{1}{\pi}
         \int\limits_0^\infty \mathrm{Re}
           \left\{      \frac{\partial G(j\omega)}{\partial \theta}
                  \left[\frac{\partial G(j\omega)}{\partial \theta}\right]^{\HT}
           \right\}
           \powerspec{u}(j\omega)
           \dd{\omega} \text{.}
\label{eq:excitation:IF}
\end{align}
The information matrix is a function of the input power spectrum $\powerspec{u}$, which is the only entity that can be used to improve the quality of the estimates. 
A common optimality criterion is D-optimality where $\det{\IF}$ is maximized~\citep{Goodwin1977}.

@@

First, we look at the band-limited $1/f$ power spectrum given by
\begin{align*}
\label{eq:excitation:pink}
\phi_u(\omega) &=\begin{cases}\frac{1}{|\omega|} &\text{if $|\omega|\in[\wn \kMin\; \wn \kMax]$,}\\
                        0           &\text{otherwise}.
            \end{cases}
\end{align*}
With this input spectrum, the power inside $\wdB$ is
\begin{align}
  \frac{1}{2\pi}
    \int\limits_{(1-\damping)\wn}^{(1+\damping)\wn}
     \frac{1}{\omega} \dd{\omega}
  = \frac{1}{2\pi}
    \ln{\frac{1+\damping}{1-\damping}},
\end{align}
which is independent of $\wn$. The robustness of the $1/f$ input has been noted previously by e.g. \cite{Rojas2007}

@@

The information matrix for systems of the type \eqref{eq:excitation:sumSys} will have a block structure commensurate with the dimension of $\theta_i$ given by
\begin{align}
\IF &= \frac{1}{\pi}
\begin{bmatrix}
  \partial G_{1,1}        &\partial G_{1,2}       &\deemph{\cdots}         &\partial G_{1,K}\\
  \partial G_{2,1}        &\partial G_{2,2}       &\deemph{\cdots}         &\partial G_{2,K}\\
  \deemph{\vdots}                  &\deemph{\vdots}                 &\deemph{\ddots}         &\deemph{\vdots}          \\
  \partial G_{K,1}        &\partial G_{K,2}       &\deemph{\cdots}         &\partial G_{K,K}
\end{bmatrix},\\
\partial G_{i,j} &= \int_0^\infty
                      \mathrm{Re}\left\{
                          \frac{\partial G_i}
                               {\partial \theta_i}
                          \frac{\partial G_j^{\HT}}
                               {\partial \theta_j}
                                \right\}
                      \phi_u(\omega)\dd\omega.
\end{align}

@@

Direct calculations for $\damping \ll 1$ give
\begin{align}
\label{eq:excitation:auto}
\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega
 = 
 \frac{\pi}
          {8\wn[1]^2}\damping^{-3} 
+ \mathcal{O}(\damping^{-1}).
\end{align}
\begin{personalNote} 
Value computed using Sage:
\begin{align*}
\frac{\pi {K}^{2}}
          {16  \sqrt{1 - \damping^2}\wn^{2} \damping^{3}} 
+
\frac{{\left(2  \damping^{2} + 1 \right)}  {K}^{2}}
         {4  \sqrt{1-\damping^{2}} \wn^{2} \damping^{2}}
+ 
\frac{{K}^{2} \arctan  
                               \left(
                                       \frac{(2  \damping^{2} - 1) \sqrt{1-\damping^{2}}}
                                                {2  (\damping^{3} - \damping)}
                              \right)}
         {8  \sqrt{1-\damping^{2}} \wn^{2}  \damping^{3}}
\end{align*}
Remark that the latter term for $\damping \to 0$ becomes
\begin{equation}
  \frac{{K}^{2} \pi / 2}
         {8  \sqrt{1-\damping^{2}} \wn^{2}  \damping^{3}}
\end{equation}
as the argument for the $\arctan$ function tends to $\infty$.
\end{personalNote}
%
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
\begin{gather}
\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right|
\qquad \qquad \qquad \qquad \qquad
\label{eq:excitation:cross1} \\
\leq
\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega \text{.}\label{eq:excitation:cross}
\end{gather}
To proceed, consider
\begin{align}
\frac{\partial G_1}{\partial \wn[1]} 
  &= \frac{2\damping j\omega\left[1+\frac{j\omega}{\wn[1]\damping}\right]}
  {\wn[1]^2\left[1+2\frac{\damping j\omega}{\wn[1]} + \left(\frac{j\omega}{\wn[1]}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\left\{0, \damping\wn[1]\right\}$. 
Hence $\left|\frac{\partial G_1}{\partial \wn[1]} \right|$ has asymptotic slope 20\unit{dB/\decade} for low frequencies, at $j\omega=\damping\wn[1]$ the slope increases to 40\unit{dB/\decade}. 
There are four poles which give a $-40$\unit{dB/\decade} roll-off after the peak. 
The shape of $\frac{\partial G_1}{\partial \wn[1]} $ is illustrated in \figref{fig:excitation:derivatives}.

>>>>

In open loop experiments, the scaled average information matrix is given by
\begin{align}
\IF & = \lim_{N\rightarrow\infty}
          \frac{1}{N}
            \sum_{t=1}^{N}
              {\left(\frac{\partial\hat{y}(t)}{\partial\theta}\right)
               \left(\frac{\partial\hat{y}(t)}{\partial\theta}\right)^{\T}},\\
\frac{\partial\hat{y}(t)}{\partial\theta}
    & =\frac{\partial G(s,\theta)}{\partial \theta}u(t).
\end{align}
By Parseval's theorem, $\IF$ can be expressed as
\begin{align}
\IF &= \frac{1}{\pi}
         \int\limits_0^\infty \RealPart
           \left\{      \frac{\partial G(j\omega)}{\partial \theta}
                  \left[\frac{\partial G(j\omega)}{\partial \theta}\right]^{\HT}
           \right\}
           \powerspec{u}(j\omega)
           \dd{\omega} \text{.}
\label{eq:excitation:IF}
\end{align}
The information matrix is a function of the input power spectrum $\powerspec{u}$, which is the only entity that can be used to improve the quality of the estimates. 
A common optimality criterion is D-optimality where $\det{\IF}$ is maximized~\citep{Goodwin1977}.

@@

First, we look at the band-limited $1/f$ power spectrum given by
\begin{align*}
\label{eq:excitation:pink}
\phi_u(\omega) &=\begin{cases}\frac{1}{|\omega|} &\text{if $|\omega|\in[\wn \kMin\; \wn \kMax]$,}\\
                        0           &\text{otherwise}.
            \end{cases}
\end{align*}
With this input spectrum, the power inside $\wdB$ is
\begin{align}
  \frac{1}{2\pi}
    \int\limits_{(1-\damping)\wn}^{(1+\damping)\wn}
     \frac{1}{\omega} \dd{\omega}
  = \frac{1}{2\pi}
    \ln{\frac{1+\damping}{1-\damping}},
\end{align}
which is independent of $\wn$. The robustness of the $1/f$ input has been noted previously by e.g. \citet{Rojas2007,Rojas2012,Goodwin2006GBO}.

@@

The information matrix for systems of the type \eqref{eq:excitation:sumSys} will have a block structure commensurate with the dimension of $\theta_i$ given by
\begin{align}
\IF &= \frac{1}{\pi}
\begin{bmatrix}
  \partial G_{1,1}        &\partial G_{1,2}       &\deemph{\cdots}         &\partial G_{1,K}\\
  \partial G_{2,1}        &\partial G_{2,2}       &\deemph{\cdots}         &\partial G_{2,K}\\
  \deemph{\vdots}                  &\deemph{\vdots}                 &\deemph{\ddots}         &\deemph{\vdots}          \\
  \partial G_{K,1}        &\partial G_{K,2}       &\deemph{\cdots}         &\partial G_{K,K}
\end{bmatrix},\\
\partial G_{i,j} &= \int_0^\infty
                      \RealPart\left\{
                          \frac{\partial G_i}
                               {\partial \theta_i}
                          \frac{\partial G_j^{\HT}}
                               {\partial \theta_j}
                                \right\}
                      \phi_u(\omega)\dd\omega.
\end{align}

@@

The $(1,1)$ element of the $\partial G_{1,1}$ entry of the information matrix is
\begin{align}
  \int_0^\infty\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega
\end{align}
which yields
\begin{equation}
\frac{\pi }
          {16  \sqrt{1 - \damping^2}\wn^{2} \damping^{3}} 
+ 
\frac{ 2\arctan  
                               \left(
                                       \frac{(2  \damping^{2} - 1) \sqrt{1-\damping^{2}}}
                                                {2  (\damping^{3} - \damping)}
                              \right)}
         {16  \sqrt{1-\damping^{2}} \wn^{2}  \damping^{3}}
+
\frac{{\left(2  \damping^{2} + 1 \right)} }
         {4  \sqrt{1-\damping^{2}} \wn^{2} \damping^{2}}
\label{eq:excitation:auto:full}
\end{equation}
after lengthy calculations.
For $\damping \ll 1$ this can be simplified to
\begin{equation}
\label{eq:excitation:auto}
\int_0^\infty\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega
 = 
 \frac{\pi}
          {8\wn[1]^2}\damping^{-3} 
+ \bigO{\damping^{-2}}
\end{equation}
since the $\arctan$ in \eqref{eq:excitation:auto:full} tends to $\frac{\pi}{2}$ for $\damping \to 0$.
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
\begin{gather}
\left|\int_0^\infty\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right|
\qquad \qquad \qquad \qquad \qquad
\label{eq:excitation:cross1} \\
\leq
\int_0^\infty\left|\RealPart\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega \text{.}\label{eq:excitation:cross}
\end{gather}
To proceed, consider
\begin{align}
\frac{\partial G_1}{\partial \wn[1]} 
  &= \frac{2\damping j\omega\left[1+\frac{j\omega}{\wn[1]\damping}\right]}
  {\wn[1]^2\left[1+2\frac{\damping j\omega}{\wn[1]} + \left(\frac{j\omega}{\wn[1]}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\left\{0, \damping\wn[1]\right\}$. 
Hence $\left|\frac{\partial G_1}{\partial \wn[1]} \right|$ has asymptotic slope 20\unit{dB/\decade} for low frequencies, at $j\omega=\damping\wn[1]$ the slope increases to 40\unit{dB/\decade}. 
There are four poles which give a $-40$\unit{dB/\decade} roll-off after the peak. 
The shape of $\frac{\partial G_1}{\partial \wn[1]} $ is illustrated in \figref{fig:excitation:derivatives}.
